---
title: "MCLPG_singlecellDA_miloR"
author: "jenssle"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Load libraries.

```{r}
library(tidyverse)
library(Seurat)
library(scCustomize)
library(miloR)
library(scran)
library(scater) 
library(patchwork)
library(cowplot)
library(SingleCellExperiment)
`%not_in%` <- purrr::negate(`%in%`)
```

Load integrated data object

```{r}
scmcl <- readRDS("/data/scMCL/scmcl_integrated.rds")
```

## DA testing with miloR

Generate the milo object.

```{r}
DefaultAssay(scmcl) <- "RNA"
# transform the Seurat object into an SingleCellExperiment object
sce.ob <- as.SingleCellExperiment(scmcl)
milo.sce <- Milo(sce.ob)
rm(sce.ob)
```

Build the KNN graph.

```{r}
set.seed(069)
milo.sce <- buildGraph(milo.sce, k =150, d = 30, reduced.dim = "SCVI")
```

Define representative neighbourhoods on KNN graph. As a rule of thumb the authors of miloR recommend an average neighbourhood size over 5 x N_samples.

```{r}
set.seed(069)
milo.sce <- makeNhoods(milo.sce, prop = 0.1, k = 150, d=30, refined = TRUE, reduced_dims = "SCVI")
plotNhoodSizeHist(milo.sce)
```

Count the cells per neighborhood.

```{r}
set.seed(069)
milo.sce <- countCells(milo.sce, 
                       meta.data= as.data.frame(colData(milo.sce)),
                       sample = "orig.ident")

head(nhoodCounts(milo.sce))
```

Build the design matrix.

```{r}
milo_design <- data.frame(colData(milo.sce))[, c("orig.ident",
                                                 "dataset",
                                                 "chemistry",
                                                 "Tissue_source",
                                                 "Timepoint")] |>
  left_join(
    read_csv("/data/scMCL/scmcl_PGprediction.csv") |>
      dplyr::select(sample_id, Cluster) |>
      dplyr::rename("orig.ident" = "sample_id"),
    by = "orig.ident"
  ) |> mutate(Cluster = case_when(grepl("ton|Con", orig.ident) ~ "normal", TRUE ~ Cluster))

milo_design <- milo_design %>%
  mutate(PG1 = case_when(Cluster == "MCL_PG1" ~ "1", TRUE ~ "0"),
         PG2 = case_when(Cluster == "MCL_PG2" ~ "1", TRUE ~ "0"), 
         PG4 = case_when(Cluster == "MCL_PG4" ~ "1", TRUE ~ "0")) %>%
  mutate(chemistry = case_when(grepl("3prime", chemistry) ~ "3prime", TRUE ~ "5prime")) |>
  distinct(orig.ident, .keep_all = TRUE)

#set the rownames
rownames(milo_design) <- milo_design$orig.ident
```

Compute the neighborhood connectivity.

```{r}
milo.sce <- calcNhoodDistance(milo.sce, d=30, reduced.dim = "SCVI")
```

Next, run the testing. We will fit the dataset as a batch factor covariate.

```{r}
da_results <- list() 

for(i in c("PG1", "PG2", "PG4")) {
  
  da_results[[i]] <- testNhoods(milo.sce, 
                         design = as.formula(paste0("~ dataset +", i)), 
                         design.df = milo_design,
                         reduced.dim = "SCVI",
                         fdr.weighting="graph-overlap"
                         )
  
}
```

Save these results.

```{r}
saveRDS(da_results, "/data/scMCL/miloR_DA_results.rds")
```

