---
title: "MCLPG_PG_inference"
author: "jenssle"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries 

```{r}
library(tidyverse)
library(MultiAssayExperiment)
library(MOFA2)
library(rhdf5)
library(cola)
library(RANN)
```

## Extract latent embedding

Using MuVI, we integrated the processed RNA and protein expression data. Load the trained object. 

```{r}
#defne´the helper function to adjust the factor names
load_muvi <- function(file) {
  model <- load_model(file,
    sort_factors = FALSE,
    remove_inactive_factors = FALSE
  )
  factors_names(model) <- h5read(file, "/factors/group_0")
  return(model)
}

muvimcl <- load_muvi("muvimcl.hdf5")
```


## Consensus graph-based clustering on this latent embedding

Generate adjacency matrix.

```{r}
t_space <- get_factors(muvimcl, factors = "all")$group_0 %>%
  as.data.frame() %>% as.matrix() %>% t()

set.seed(42)

snn <- RANN::nn2(t(t_space), k=15)$nn.idx


adjacency_matrix <- matrix(0L, ncol(t_space), ncol(t_space))
rownames(adjacency_matrix) <- colnames(adjacency_matrix) <- colnames(t_space)

for(i in 1:ncol(t_space)) {
    adjacency_matrix[i,colnames(t_space)[snn[i,]]] <- 1L
}

#check that rows add to k
sum(adjacency_matrix[1,]) == 15
table(apply(adjacency_matrix, 1, sum))
```

Iterative Leiden-alg based community detection

```{r}
#Retrieve names for labeling of the resulting leiden-communities x samples matrix
Latent_space_prots1 <- get_factors(muvimcl, factors = "all")$group_0 %>%
  as.data.frame() %>% tibble::rownames_to_column("Prot_id")

ident_global <- data.frame(Prot_id = Latent_space_prots1$Prot_id)

for(i in 1:1000) {

set.seed(sample(1:1000,1))
partition_new <- leiden(adjacency_matrix, 
                        partition_type = c("RBConfigurationVertexPartition"), 
                        resolution_parameter = 1, 
                        n_iterations = -1L, 
                        seed = sample(1:1000,1)
                        )

ident_partition <- cbind(Latent_space_prots1$Prot_id, partition_new) %>% as.data.frame()

colnames(ident_partition) <- c("Prot_id", i)

ident_global <- left_join(ident_global, ident_partition, by = "Prot_id")

}

#get rownames and transpose the resulting df
ident_global <- ident_global %>%
  column_to_rownames("Prot_id") %>%
  t() %>%
  as.data.frame() 

#mutate all variables to numeric
ident_global_m <- mutate_all(ident_global, function(x) as.numeric(as.character(x)))
```

Now, register the "take all" method for top values.

```{r}
register_top_value_methods(
 
  all = function(mat) {
    
    all <- c(runif(n = nrow(mat), min = 1, max = 1))
 
  }
 
)
```

Second, register hclust on the basis of hamming distance.

```{r}
register_partition_methods(
    hclust_hamming = function(mat, k) {
      
      library(tidyverse)
      library(e1071)
      
      mat_fac <- mutate_all(as.data.frame(mat),as.factor)
        
      
      ham_dist_fac <- e1071::hamming.distance(as.matrix(t(mat_fac)))
      
      res_clust <- cutree(hclust(as.dist(ham_dist_fac)), k)
      
      res_clust
    }
)
```

Run the iterative clustering of the generated leiden matrix.

```{r}
res <- consensus_partition(
  as.matrix(ident_global_m),
  top_value_method = c("all"),
  partition_method = c("hclust_hamming"),
  mc.cores = 30,
  top_n = nrow(ident_global_m),
  max_k = max(ident_global_m),
  scale_rows = FALSE
  )
```

Collect the best k suggestion.

```{r}
suggest_best_k(res)

select_partition_number(res)
```

Extract the MCL-PG final assignment.

```{r}
res_class <- get_classes(res, k=8) %>%
  rownames_to_column("dbnr")
```

